Amplicon sequence variants artificially split bacterial genomes into separate clusters

### 2021-02-26 submission/manuscript.tex
Fifth round of revisions

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {}%
  {\par}

@@

Amplicon sequencing variants (ASVs) have been proposed as an alternative
to operational taxonomic units (OTUs) for analyzing microbial
communities. ASVs have grown in popularity, in part, because of a desire
to reflect a more refined level of taxonomy since they do not cluster
sequences based on a distance-based threshold. However, ASVs and the use
of overly narrow thresholds to identify OTUs increase the risk of
splitting a single genome into separate clusters. To assess this risk, I
analyzed the intragenomic variation of 16S rRNA genes from the bacterial
genomes represented in a \emph{rrn} copy number database, which
contained \textbf{20,427} genomes from \textbf{5,972} species. As the
number of copies of the 16S rRNA gene increased in a genome, the number
of ASVs also increased. There was an average of \textbf{0.58} ASVs per
copy of the 16S rRNA gene for full length 16S rRNA genes. It was
necessary to use a distance threshold of \textbf{5.25}\% to cluster full
length ASVs from the same genome into a single OTU with 95\% confidence
for genomes with 7 copies of the 16S rRNA, such as \emph{E. coli}. This
research highlights the risk of splitting a single bacterial genome into
separate clusters when ASVs are used to analyze 16S rRNA gene sequence
data. Although there is also a risk of clustering ASVs from different
species into the same OTU when using broad distance thresholds, those
risks are of less concern than artificially splitting a genome into
separate ASVs and OTUs.

@@

16S rRNA gene sequencing is a powerful technique for describing and
comparing microbial communities (1). Efforts to link 16S rRNA gene
sequences to taxonomic levels based on distance thresholds date to at
least the 1990s. The distance-based threshold that was developed and is
now widely used was based on DNA-DNA hybridization approaches that are
not as precise as genome sequencing (2, 3). Instead, genome sequencing
technologies have suggested that the widely used 3\% distance threshold
to operationally define bacterial taxa is too coarse (4--6). As an
alternative to OTUs, amplicon sequencing variants (ASVs) have been
proposed as a way to adopt the thresholds suggested by genome sequencing
to microbial community analysis using 16S rRNA gene sequences (7--10).
Approaches for identifying ASVs do not cluster sequences based on a
distance-based threshold (11). Proponents of ASVs are largely
dissmissive of concerns that most bacterial genomes have more than one
copy of the \emph{rrn} operon and that those copies are not identical
(12, 13). Yet, using too fine a threshold to identify OTUs creates the
risk of splitting a single genome into multiple clusters. Conversely,
using too broad of a threshold to define OTUs creates the risk of
clustering together bacterial species into the same OTU. An example of
both is seen in the comparison of \emph{Staphylococcus aureus} (NCTC
8325) and \emph{S. epidermidis} (ATCC 12228) where each genome has 5
copies of the 16S rRNA gene. Each of the 10 copies of the 16S rRNA gene
in these two genomes is distinct and represent 10 ASVs. Conversely, if
the copies were clustered using a 3\% distance threshold, then all 10
ASVs would cluster into the same OTU. The goal of this study was to
quantify the tradeoff of splitting a single genome into multiple
clusters and the risk of clustering different bacterial species into the
same cluster when using ASVs and various OTU definitions.

To investigate the variation in the number of copies of the 16S rRNA
gene per genome and the intragenomic variation among copies of the 16S
rRNA gene, I obtained 16S rRNA sequences from the \emph{rrn} copy number
database (\emph{rrn}DB)(14). Among the \textbf{5,972} species
represented in the \emph{rrn}DB there were \textbf{20,427} genomes. The
median \emph{rrn} copy number per species ranged between \textbf{1}
(e.g., \textbf{\emph{Mycobacterium tuberculosis}}) and \textbf{19}
(\textbf{\emph{Metabacillus litoralis}}). As \textbf{rrn} copy number
for a genome increased, the number of variants of the 16S rRNA gene in
each genome also increased. On average, there were \textbf{0.58}
variants per copy of the full length 16S rRNA gene and an average of
\textbf{0.32}, \textbf{0.25}, and \textbf{0.27} variants when
considering the V3-V4, V4, and V4-V5 regions of the gene, respectively.
Although a species tended to have a consistent number of 16S rRNA gene
copies per genome, the number of total variants increased with the
number of genomes that were sampled (\textbf{Figure 1}). For example,
the \textbf{271} genome accessions of \textbf{\emph{Mycobacterium
tuberculosis}} in the \emph{rrn}DB each had \textbf{1} copy of the gene
per genome. However, across those accessions, there were \textbf{17}
versions of the gene. An \emph{E. coli} genome typically had \textbf{7}
copies of the 16S rRNA gene with a median of \textbf{5} distinct full
length ASVs per genome (intraquartile range between \textbf{3} and
\textbf{6}). Across the \textbf{1,390} \emph{E. coli} genomes in the
\emph{rrn}DB, there were \textbf{1,402} versions of the gene. These
observations highlight the risk of selecting a threshold for defining
clusters that is too narrow because it is possible to split a single
genome into multiple clusters.

A method to avoid splitting a single genome into multiple clusters is to
cluster 16S rRNA gene sequences together based on their distances
between each other. Therefore, I assessed the impact of the distance
threshold used to define clusters of 16S rRNA genes on the propensity to
split a genome into separate clusters. To control for uneven
representation of genomes across species, I randomly selected one genome
from each species for the following analyses and repeated each
randomization 100 times. I observed that as the \emph{rrn} copy number
increased, the distance threshold required to reduce the ASVs in each
genome to a single OTU increased (Figure 1). Among species with 7 copies
of the \emph{rrn} operon (e.g., \emph{E. coli}), a distance threshold of
\textbf{5.25}\% was required to reduce full length ASVs into a single
OTU for 95\% of the species (Figure 2). Similarly, thresholds of
\textbf{5.25}, \textbf{2.50}, and \textbf{3.75}\% were required for the
V3-V4, V4, and V4-V5 regions, respectively. But, if a 3\% distance
threshold was used, then ASVs from genomes containing fewer than
\textbf{10}, \textbf{10}, \textbf{8}, and \textbf{10} copies of the
\emph{rrn} operon would reliably be clustered into a single OTU for ASVs
from the V1-V9, V3-V4, V4, and V4-V5 regions, respectively (Figure 2).
Consequently, these results demonstrate that broad thresholds must be
used to avoid splitting different operons from the same genome into
separate clusters. At broad thresholds, 16S rRNA gene sequences from
multiple species could be clustered into the same ASV or OTU
(\textbf{Figure 2}). Using ASVs, \textbf{4.11}\% of the ASVs contained
sequences from multiple species when considering full length sequences
and \textbf{10.87}, \textbf{16.20}, and \textbf{13.06}\% when
considering the V3-V4, V4, and V4-V5 regions, respectively. At the
commonly used 3\% threshold for defining OTUs, \textbf{27.41}\% of the
OTUs contained 16S rRNA gene sequences from multiple species when
considering full length sequences and \textbf{31.71}, \textbf{34.31},
and \textbf{34.78}\% when considering the V3-V4, V4, and V4-V5 regions,
respectively. Considering that species designations are unevenly applied
and reflect multiple human-imposed biases, the risk of splitting a
genome into multiple OTUs is more problematic than clustering species
together. Therefore, larger thresholds are advisable.

The results of this analysis demonstrate that there is a significant
risk of splitting a single genome into multiple clusters if using ASVs
or too fine of a threshold to define OTUs. An ongoing problem for
amplicon-based studies is defining a meaningful taxonomic unit (11, 15,
16). Since there is no consensus for a biologicaly definition of a
bacterial species (17--19), microbiologists must accept that how
bacterial species are named is biased and that taxonomic rules are not
applied in a consistent manner (e.g., (19, 20)). This makes it
impossible to fit a distance threshold to define an OTU definition that
matches a set of species names (21). Furthermore, the 16S rRNA gene does
not evolve at the same rate across all bacterial lineages (15), which
limits the biological interpretation of a common OTU definition. A
distance-based definition of a taxonomic unit based on 16S rRNA gene or
full genome sequences is, at best, operational and not grounded in
biological theory (15, 22--24). There is general agreement in bacterial
systematics that to classify an organism to a bacterial species,
phenotypic and genome sequence data are needed (17--20). A short
sequence from a bacterial genome simply cannot differentiate between
species. Moreover, it is difficult to defend a clustering approach that
would split a single genome into multiple taxonomic units. It is not
biologically plausible to entertain the possibility that parts of a
genome would have different ecologies. Although there are multiple
reasons that proponents favor ASVs, the significant risk of artificially
splitting genomes into separate clusters is too high to warrant their
use.

@@

\textbf{(iii) Clustering sequences into OTUs.} Pairwise distances
between sequences were calculated using the dist.seqs command from
mothur. The OptiClust algorithm as implemented in mothur was used to
assign 16S rRNA gene sequences to OTUs (27). Distance thresholds between
0.25 and 10.0\% in 0.25 percentage point increments were used to assign
sequences to OTUs.

\textbf{(iv) Controlling for uneven sampling of genomes by species.}
Because of the uneven distribution of genome sequences across species,
for the analysis of splitting genomes and clustering ASVs from different
species, I randomly selected one genome from each species. The random
selection was repeated 100 times. Analyses based on this randomization
reported the median of the 100 randomizations. The intraquartile range
between randomizations was less than \textbf{0.0024}. Because the range
was so small, the confidence intervals were more narrow than the
thickness of the lines in Figures 1 and 2 and were not included.

@@

\textbf{(vi) Note on usage of ASV, OTU, and cluster.} I used ``ASV'' to
denote the cluster of true 16S rRNA gene sequences that were identical
to each other and ``OTU'' to denote the product of distance-based
clustering of sequences. Although ASVs do represent a type of
operational defition of a taxonomic unit and can be thought of as an OTU
formed using a distance of zero, proponents of the ASV approach prefer
to avoid the term OTU given the long history of OTUs being formed by
distance-based clustering
(\url{https://github.com/benjjneb/dada2/issues/62}; accessed
2021-02-19). For this reason, when an ASV splits a genome into different
units, those units were called clusters rather than OTUs.

@@

\textbf{Figure 2. As the distance threshold used to define an OTU
increased, the fraction of genomes split into separate OTUs decreased
while the fraction of ASVs and OTUs representing multiple species
increased.} These data represent the median fractions for both
measurements across 100 randomizations. In each randomization, one
genome was sampled from each species.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\raggedright
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {}%
  {\par}

@@

Amplicon sequencing variants (ASVs) have been proposed as an alternative
to operational taxonomic units (OTUs) for analyzing microbial
communities. ASVs have grown in popularity, in part, because of a desire
to reflect a more refined level of taxonomy since they do not cluster
sequences based on a distance-based threshold. However, ASVs and the use
of overly narrow thresholds to identify OTUs increase the risk of
splitting a single genome into separate clusters. To assess this risk, I
analyzed the intragenomic variation of 16S rRNA genes from the bacterial
genomes represented in a \emph{rrn} copy number database, which
contained \textbf{20,427} genomes from \textbf{5,972} species. As the
number of copies of the 16S rRNA gene increased in a genome, the number
of ASVs also increased. There was an average of \textbf{0.58} ASVs per
copy of the 16S rRNA gene for full length 16S rRNA genes. It was
necessary to use a distance threshold of \textbf{5.2}\% to cluster full
length ASVs from the same genome into a single OTU with 95\% confidence
for genomes with 7 copies of the 16S rRNA, such as \emph{E. coli}. This
research highlights the risk of splitting a single bacterial genome into
separate clusters when ASVs are used to analyze 16S rRNA gene sequence
data. Although there is also a risk of clustering ASVs from different
species into the same OTU when using broad distance thresholds, those
risks are of less concern than artificially splitting a genome into
separate ASVs and OTUs.

@@

16S rRNA gene sequencing is a powerful technique for describing and
comparing microbial communities (1). Efforts to link 16S rRNA gene
sequences to taxonomic levels based on distance thresholds date to at
least the 1990s. The distance-based threshold that was developed and is
now widely used was based on DNA-DNA hybridization approaches that are
not as precise as genome sequencing (2, 3). Instead, genome sequencing
technologies have suggested that the widely used 3\% distance threshold
to operationally define bacterial taxa is too coarse (4--6). As an
alternative to OTUs, amplicon sequencing variants (ASVs) have been
proposed as a way to adopt the thresholds suggested by genome sequencing
to microbial community analysis using 16S rRNA gene sequences (7--10).
Approaches for identifying ASVs do not cluster sequences based on a
distance-based threshold (11). Proponents of ASVs are largely
dissmissive of concerns that most bacterial genomes have more than one
copy of the \emph{rrn} operon and that those copies are not identical
(12, 13). Yet, ASVs and using too fine a threshold to identify OTUs
could split a single genome into multiple clusters. Conversely, using
too broad of a threshold to define OTUs could cluster together multiple
bacterial species into the same OTU. An example of both is seen in the
comparison of \emph{Staphylococcus aureus} (NCTC 8325) and \emph{S.
epidermidis} (ATCC 12228) where each genome has 5 copies of the 16S rRNA
gene. Each of the 10 copies of the 16S rRNA gene in these two genomes is
distinct and represent 10 ASVs. Conversely, if the copies were clustered
using a 3\% distance threshold, then all 10 ASVs would cluster into the
same OTU. The goal of this study was to quantify the tradeoff of
splitting a single genome into multiple clusters and the risk of
clustering different bacterial species into the same cluster when using
ASVs and various OTU definitions.

To investigate the variation in the number of copies of the 16S rRNA
gene per genome and the intragenomic variation among copies of the 16S
rRNA gene, I obtained 16S rRNA sequences from the \emph{rrn} copy number
database (\emph{rrn}DB)(14). Among the \textbf{5,972} species
represented in the \emph{rrn}DB there were \textbf{20,427} genomes. The
median \emph{rrn} copy number per species ranged between \textbf{1}
(e.g., \textbf{\emph{Mycobacterium tuberculosis}}) and \textbf{19}
(\textbf{\emph{Metabacillus litoralis}}). As \emph{rrn} copy number for
a genome increased, the number of variants of the 16S rRNA gene in each
genome also increased. On average, there were \textbf{0.58} variants per
copy of the full length 16S rRNA gene and an average of \textbf{0.32},
\textbf{0.25}, and \textbf{0.27} variants when considering the V3-V4,
V4, and V4-V5 regions of the gene, respectively. Although a species
tended to have a consistent number of 16S rRNA gene copies per genome,
the number of total variants increased with the number of genomes that
were sampled (\textbf{Figure 1}). For example, the \textbf{271} genome
accessions of \textbf{\emph{Mycobacterium tuberculosis}} in the
\emph{rrn}DB each had \textbf{1} copy of the gene per genome. However,
across those accessions, there were \textbf{17} versions of the gene. An
\emph{E. coli} genome typically had \textbf{7} copies of the 16S rRNA
gene with a median of \textbf{5} distinct full length ASVs per genome
(intraquartile range between \textbf{3} and \textbf{6}). Across the
\textbf{1,390} \emph{E. coli} genomes in the \emph{rrn}DB, there were
\textbf{1,402} versions of the gene. These observations highlight the
risk of selecting a threshold for defining clusters that is too narrow
because it is possible to split a single genome into multiple clusters.

A method to avoid splitting a single genome into multiple clusters is to
cluster 16S rRNA gene sequences together based on their distances
between each other. Therefore, I assessed the impact of the distance
threshold used to define clusters of 16S rRNA genes on the propensity to
split a genome into separate clusters. To control for uneven
representation of genomes across species, I randomly selected one genome
from each species and repeated each randomization 100 times. I observed
that as the \emph{rrn} copy number increased, the distance threshold
required to reduce the ASVs in each genome to a single OTU increased
(Figure 1). Among species with 7 copies of the \emph{rrn} operon (e.g.,
\emph{E. coli}), a distance threshold of \textbf{5.25}\% was required to
reduce full length ASVs into a single OTU for 95\% of the species
(Figure 1). Similarly, thresholds of \textbf{5.25}, \textbf{2.50}, and
\textbf{3.75}\% were required for the V3-V4, V4, and V4-V5 regions,
respectively. But, if a 3\% distance threshold was used, then ASVs from
genomes containing fewer than \textbf{10}, \textbf{10}, \textbf{8}, and
\textbf{10} copies of the \emph{rrn} operon would reliably be clustered
into a single OTU for ASVs from the V1-V9, V3-V4, V4, and V4-V5 regions,
respectively (Figure 1). Consequently, these results demonstrate that
broad thresholds must be used to avoid splitting different operons from
the same genome into separate clusters.

At broad thresholds, 16S rRNA gene sequences from multiple species could
be clustered into the same ASV or OTU (Figure 2). For this analysis, I
again randomly selected one genome from each species to control for
uneven representation of genomes across species and repeated each
randomization 100 times. Without using distance-based thresholds,
\textbf{4.1}\% of the ASVs contained sequences from multiple species
when considering full length sequences and \textbf{10.9}, \textbf{16.2},
and \textbf{13.1}\% when considering the V3-V4, V4, and V4-V5 regions,
respectively. At the commonly used 3\% threshold for defining OTUs,
\textbf{27.4}\% of the OTUs contained 16S rRNA gene sequences from
multiple species when considering full length sequences and
\textbf{31.7}, \textbf{34.3}, and \textbf{34.8}\% when considering the
V3-V4, V4, and V4-V5 regions, respectively. Considering that species
designations are unevenly applied and reflect multiple human-imposed
biases, the risk of splitting a genome into multiple OTUs is more
problematic than clustering species together. Therefore, larger
thresholds are advisable.

The results of this analysis demonstrate that there is a significant
risk of splitting a single genome into multiple clusters if using ASVs
or too fine of a threshold to define OTUs. An ongoing problem for
amplicon-based studies is defining a meaningful taxonomic unit (11, 15,
16). Since there is no consensus for a biologicaly definition of a
bacterial species (17--19), microbiologists must accept that how
bacterial species are named is biased and that taxonomic rules are not
applied in a consistent manner (e.g., (19, 20)). This makes it
impossible to fit a distance threshold to define an OTU definition that
matches a set of species names (21). Furthermore, the 16S rRNA gene does
not evolve at the same rate across all bacterial lineages (15), which
limits the biological interpretation of a common OTU definition. A
distance-based definition of a taxonomic unit based on 16S rRNA gene or
full genome sequences is, at best, operational and not grounded in
biological theory (15, 22--24). There is general agreement in bacterial
systematics that to classify an organism to a bacterial species,
phenotypic and genome sequence data are needed (17--20). A short
sequence from a bacterial genome simply cannot differentiate between
species. Moreover, it is difficult to defend a clustering threshold that
would split a single genome into multiple taxonomic units. It is not
biologically plausible to entertain the possibility that different
\emph{rrn} operons from the same genome would have different ecologies.
Although there are multiple reasons that proponents favor ASVs, the
significant risk of artificially splitting genomes into separate
clusters is too high to warrant their use.

@@

\textbf{(iii) Clustering sequences into OTUs.} Pairwise distances
between sequences were calculated using the dist.seqs command from
mothur. The OptiClust algorithm, as implemented in mothur, was used to
assign 16S rRNA gene sequences to OTUs (27). Distance thresholds between
0.25 and 10.00\% in 0.25 percentage point increments were used to assign
sequences to OTUs.

\textbf{(iv) Controlling for uneven sampling of genomes by species.}
Because of the uneven distribution of genome sequences across species,
for the analysis of splitting genomes and clustering ASVs from different
species (Figures 1 and 2), I randomly selected one genome from each
species. The random selection was repeated 100 times. Analyses based on
this randomization reported the median of the 100 randomizations. The
intraquartile range between randomizations was less than
\textbf{0.0024}. Because the range was so small, the confidence
intervals were more narrow than the thickness of the lines in Figures 1
and 2 and were not included.

@@

\textbf{(vi) Note on usage of ASV, OTU, and cluster.} I used ``ASV'' to
denote the cluster of true 16S rRNA gene sequences that were identical
to each other and ``OTU'' to denote the product of distance-based
clustering of sequences. Although ASVs do represent a type of
operational defition of a taxonomic unit and can be thought of as an OTU
formed using a distance of zero, proponents of the ASV approach prefer
to avoid the term OTU given the long history of OTUs being formed by
distance-based clustering
(\url{https://github.com/benjjneb/dada2/issues/62}; accessed
2021-02-26). For this reason, when an ASV split a genome into different
units, those units were called clusters rather than OTUs.

@@

\textbf{Figure 2. As the distance threshold used to define an OTU
increased, the percentage of ASVs and OTUs representing multiple species
increased.} These data represent the median fractions for both
measurements across 100 randomizations. In each randomization, one
genome was sampled from each species.
